### 2013-08-29 thesis/src/chapters/chapter-2.tex
Added horizontal scaling stub

What makes scaling application particularly interesting are the benefits offered by a cloud computing, especially the illusion of a virtually infinite computing infrastructure \cite{VaRoBu11}. Making use of virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner.

@@

The rest of this chapter elaborates in details about horizontal and vertical scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.

\section{Horizontal scaling}

As it was outlined in a previous section, vital element of horizontal scaling is adding additional resources. In context of an application execution environment, node can be represented by an application server, database or other resources that affect application performance such as network interface \cite{VaRoBu11}. Dynamic capabilities of a cloud computing platform offer a couple of ways this can be done:
\begin{itemize}
	\item virtual machine replication
	\item network slicing
	\item database replication
\end{itemize}

jebnac mapowanie na warstwy clouda - infr, platform, app

\subsection{Virtual machine replication}

\subsection{Network slicing}

\subsection{Database replication}

\section{Vertical scaling}

>>>>

What makes scaling application particularly interesting are the benefits offered by a cloud computing, especially the illusion of a virtually infinite computing infrastructure \cite{VaRoBu11}. Making use of virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner. Although, cloud computing offers additional scaling capabilities, it increases solution complexity since they operate in different layers: server, platform, network as stated in \cite{VaRoBu11}.

@@

The rest of this chapter elaborates in detail about server, platform and network scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.

\section{Server level}

\subsection{Horizontal scaling}
As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system. As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion. As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later. While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack, OpenNebula, OpenStack, CloudStack or Eucalyptus to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done. Table 1 describes in details different capabilities of commonly used virtualization technologies.

Provisioning new server is only a first step in scaling an application, it is required to configure some load balancing mechanism to make user of this additional node.

\subsection{Vertical scaling}

\section{Platform level}

\section{Network level}
